PMV: Which formulation should you use the ISO 7730:2005 or ASHRAE 55:2020?

### 2023-08-29 Manuscript/src/sections/results.tex
paper: reviewing Stefano's comments

%! Author = sbbfti
%! Date = 10/06/2020

\section{Results}\label{sec:results}

\subsection{Comfort DB overview}\label{subsec:comfort-db-overview}

A total of \var{entries_db_used} entries met the inclusion criteria listed in the Methodology Section.
The distribution of the six input variables used to calculate the PMV indices is depicted in Figure~\ref{fig:dist_input_data}.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/dist_input_data}
    \caption{Distribution of the input variables used to calculate the \ac{pmv} values.
    The data are shown using boxen-plots (letter-value plots).
    They depict the median as the centerline and each successive level outward contains half of the remaining data.}
    \label{fig:dist_input_data}
\end{figure}

Figure~\ref{fig:dist_other_data} shows the distribution of the age, height, weight, and running mean outdoor temperature grouped by sex.
Less than half (\num{23300}) of the total entries had information about the participant's sex.
Out of those data were almost equally distributed among male \qty{52}{\percent} and females.
Ages are not normally distributed, the same is true for the running mean outdoor temperature.
About half of the participants (\qty{52}{\percent}) were aged between \num{20} and \num{35} years old.
While \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
Mens were significantly heavier and taller than females.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/dist_other_data}
    \caption{Distribution of age, height, weight, and running mean outdoor temperature.
    The data are grouped by sex.}
    \label{fig:dist_other_data}
\end{figure}

Figure~\ref{fig:bar_plot_tp_by_ts} shows on the left the percentage of thermal preference votes grouped by each \ac{tsv}.
The left Figure also reports the number of points for each \ac{tsv}.
Only \var{entries_with_tp} that had information about the thermal preference.
Thus, on the right we show a bar chart depicting the distribution of all the \ac{tsv} votes, we have used in the analysis.
The \gls{db2} in respect to the \ac{tsv} is unbalanced.
Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
In the \gls{db2} more than \qty{50}{\percent} of participants who were either `slightly warm' or `slightly cool' wanted to be `cooler' or `warmer', respectively.
However, in thermal comfort research it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable.
This finding challenges the above assumption.
Moreover, these results provide additional evidence that the use of the thermal preference scale is better suited to determine how people perceive their thermal environment.
The thermal sensation vote clearly allows to determine whether participants would take action or not to modify their current thermal environment.
The \ac{tsv}, on the contrary, does not allow to determine which actions should be taken to increase thermal comfort conditions for occupants.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/bar_plot_tp_by_ts}
    \caption{The left Figure shows the percentage of thermal preference votes for each thermal sensation vote.
    The numbers on the right side of each bar show the number of points for each \ac{tsv}.
    The right figure shows the total number of data points grouped by \ac{tsv}.
    Each bin in the Figure of the right has more data points than in the Figure on the left since information about the occupants thermal preference were not always available.}
    \label{fig:bar_plot_tp_by_ts}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
    \caption{}
    \label{fig:bubble_models_vs_tsv}
\end{figure}

>>>>

%! Author = sbbfti
%! Date = 10/06/2020
\section{Results and Discussion}\label{sec:results}

\subsection{\ac{db2} overview}\label{subsec:comfort-db-overview}
While the \ac{db2} contains \var{entries_db_all} data points, only \var{entries_db_used} entries met the inclusion criteria listed in the Methodology section.
The distribution of the six input variables used to calculate the PMV indices is visualized in Figure~\ref{fig:dist_input_data}.
The 2.5th, 50th (median), and 97.5th percentiles are shown in blue.
% todo: remove the gridlines or marker them very short; remove outliers or make the line very thin; keep one decimal for clo values; remove decimal point for RH; annotation font should be the same as the label and same size, change only the color; You can use black or blue (HEX: 003262, RGB (0, 50, 98)) for the chart; one decimal for met not two; the chart is a bit dense of information
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/dist_input_data}
    \caption{Distribution of the input variables used to calculate the \ac{pmv} and \ac{pmv-ce} values.
    The text in blue is showing the 2.5th, 50th (median), and 97.5th percentiles.
    The data are shown using boxen plots (letter-value plots).
    They depict the median as the center-line and each successive level outward contains half of the remaining data.}
    \label{fig:dist_input_data}
\end{figure*}
The input data distributions are far from uniformity (equal density) across the Standards' applicability ranges.
We were not able to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} formulations for inputs that were lower than the 2.5th and 97.5th percentiles, since we only had a limited number of data points beyond those values.
For example, only 2.5\% of the available dataset had a \ac{tdb} $\leq$ than \var{ta_95_perc_min}, this may consequently be an issue since both thermal comfort Standards' applicability limits extend to \qty{10}{\celsius}.
This can be explained by the fact that only a few buildings in the world operate at \ac{tdb} lower than this value.
This is because an indoor \ac{tdb} lower than \var{ta_95_perc_min} would be deemed to be unsatisfactory by a great number of occupants~\cite{iso7730}.
This lack of data in the \ac{db2} is particularly relevant for \ac{v}.
The \ac{pmv} and \ac{pmv-ce} only differ when the value of \ac{v} $\geq$ \qty{0.1}{\m\per\s}, however, the median value in the \ac{db2} was \qty{0.08}{\m\per\s} suggesting that most of the data points were collected in buildings with low air movement.
Consequently, we could not reliably test the accuracy of the \ac{pmv} formulations for values of \ac{v} $\geq$ \var{v_95_perc_max}.

Figure~\ref{fig:dist_other_data} shows the distribution of age, height, weight, and running mean outdoor temperature grouped by sex.
Less than half (\num{23300}) of the total entries have information about the participant's sex and are almost equally distributed among males \qty{52}{\percent} and females.
Men were significantly heavier and taller than females.
Ages are not normally distributed, the same is true for the running mean outdoor temperature.
About half of the participants (\qty{52}{\percent}) were aged between \num{20} and \num{35} years old, and only \qty{2.5}{\percent} were older than 60.
This limits our analysis to young adults and does not allow us to determine the accuracy of the \ac{pmv} and \ac{pmv-ce} models in predicting thermal sensation for children or older adults.
The \ac{db2} does not contain information about the health status of the participants.
Approximately \qty{56}{\percent} of the running mean outdoor temperature values where between \qtyrange{10}{25}{\celsius}.
% todo: check that the colours are distinguishable for color blind and change them since I am using the same color palette to differentiate PMV and PMVCE; same comments as previous chart; use a boxen plot and have 4 for men and 4 for women. Tormt is missing the unit; age unot should be year. This figure can be merged with figure 2 because it is a visualization of the database. I would use blue (HEX: 003262, RGB (0, 50, 98)) and gold (HEX: FDB515, RGB (253, 181, 21)) as color because blue and green could be confounded by colorblind people. To be more consisten in the paper, I would use the letter value plots for this and the figure in section 3.3.2
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/dist_other_data}
    \caption{Distribution of age, height, weight, and running mean outdoor temperature.
    The data are grouped by sex.
    The text in blue is showing the 2.5th, 50th (median), and 97.5th percentiles.}
    \label{fig:dist_other_data}
\end{figure*}
The percentage of \ac{tpv} grouped by each \ac{tsv} is shown on the left side of Figure~\ref{fig:bar_plot_tp_by_ts}.
Only \var{entries_with_tp} entries had information about both \ac{tsv} and \ac{tpv}.
Thus, on the right, we show a bar chart depicting the distribution of all the \ac{tsv} votes, we analyzed in the paper.
The \ac{tsv} is unbalanced. 
Approximately \var{perc_tsv_neutral} of all the entries have a \ac{tsv} of `Neutral'.
While less than \var{perc_tsv_hot} of the total sample of participants reported to be either `Hot' or `Cold'.
% todo I really do not like this color, they are misleading. If someone want to be "warmer", then it should be shown as blue. I suggest to use the color pallet (figure 3) that Micheal used in the BCA paper. 
% todo TSV report words not numbers.
% todo TSV chart the color here should not match the one on the thermal preference scale given that it is a different question. For example, a TSV=3 is red, like people that wants to be warmer, meaning that they are cold. Even with the inversion of the color palette, we cannot imply that TSV=3 mease wanted to be cooler. What about TSV=1, maybe people want no changer or cooler or warmer. Difficult to assume. Better to use two independet color schemes. 
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bar_plot_tp_by_ts}
    \caption{The left Figure shows the percentage of \ac{tpv} for each thermal sensation vote.
    The numbers on the right side of each bar show the number of points for each \ac{tsv}.
    The right figure shows the total number of data points grouped by \ac{tsv}.
    Each bin in the right figure has more data points than in the left figure since less data on \ac{tpv} were available.}
    \label{fig:bar_plot_tp_by_ts}
\end{figure*}

In thermal comfort research, it is generally assumed that people who are `slightly warm' or `slightly cool' are thermally comfortable.
However, in the \ac{db2} more than \qty{50}{\percent} of participants who were either `slightly warm' or `slightly cool' wanted to be `cooler' or `warmer', respectively.
This finding challenges the above assumption. 
Assuming that `slightly warm' or `slightly cool' is comfortable leads to under-counting thermal discomfort.
This result highlights one of the major limitations of the \ac{tsv} scale and suggests that the thermal preference scale is better suited to determine how people perceive their thermal environment.
The \ac{tpv} more clearly allows an HVAC system to decide whether to change or not the thermal environment. 
On the contrary, the \ac{tsv} is ambiguous and does not allow determining if an action should be taken to increase the participant's thermal comfort.

\subsection{Comparison of PMV accuracy in predicting thermal sensation}\label{subsec:model-accuracy-comparison-in-predicting-thermal-sensation}
The \ac{pmv} was developed with the primary aim of predicting \ac{tsv}, consequently, we are here comparing the prediction accuracy of the two formulations.
% todo Use the words, not the numbers. Please use the color pallet from Toby's paper. It is hard to notice that these are two distint graph because the legend is moving accross. I suggest to move the legend above and have a thin grey line to separate the two group, another option would be to add a light grey background for the two PMV and show that they are distinct. 
% todo Move the percentage, on top and move it horizontally.
% todo remove the white space between the boxes (horizontal lines)
% todo TSV use words not numbers
% todo Move this also above the figure, over each column. I would also add the term. Overall accuracy  PMV = 33%
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bar_stacked_model_accuracy}
    \caption{The calculated \ac{pmv} value is grouped by the \ac{tsv} value reported by the participants. 
    The number of data points in each bin of the two charts is the same and this allows us to compare the accuracy of both models.
    The overall prediction accuracy is shown above the chart while the model accuracy of predicting participants grouped by their \ac{tsv} is overlayed over each bar. 
    A random model, randomly selecting a thermal sensation vote would have an overall accuracy of \qty{14}{\percent}}
    \label{fig:bar_stacked_model_accuracy}
\end{figure*}
Being \ac{tsv} the value reported by participants (i.e., ground truth), in Figure~\ref{fig:bar_stacked_model_accuracy} we grouped participants' responses by \ac{tsv} and reported the simple accuracy of the different \ac{pmv} formulations over each sub-figure.
Both \ac{pmv} formulations underestimated the conditions at which participants are thermally dissatisfied with their environment.
The overall accuracy of the \ac{pmv} and \ac{pmv-ce} were \qty{33}{\percent} and \qty{34}{\percent}, respectively.
Being the classification problem unbalanced, the simple accuracy is not an optimal metric to rate the model prediction accuracy since a default strategy of guessing the majority class would lead to a high overall model accuracy.
A model that would have always predicted `neutral' would have achieved an overall accuracy of \var{perc_tsv_neutral}, the number of participants who voted \ac{tsv}=0.
For this reason, in Table~\ref{tab:f1}, we report the F1-scores for all the models.
The F1-macro score, which is free from label imbalance, depicts that the accuracy of both models across all metrics is marginally better than random guessing (i.e., \qty{14}{\percent}).
% todo in the table I should also report the F1 scores if I filter out some of the TSV. The PMV model should only be used in the range -2 and 2 so at least we should report also the results in this range.
\begin{table}[htb!]
    \centering
    \input{tables/f1.tex}
    \caption{F1-score for the \ac{pmv} and \ac{pmv-ce} models.}
    \label{tab:f1}
\end{table}

The \ac{pmv-ce} was specifically developed to more accurately estimate latent and sensible heat losses from the skin to the environment when air movement is used.
However, our results show that the \ac{pmv-ce} is less accurate than the \ac{pmv} in predicting people who were `slightly warm'.
Both models had an accuracy lower than \qty{1}{\percent} when predicting either `warm', `hot', or `cold'.
This value is significantly lower and worse than random guessing, this is very concerning because this means that the models are actively misleading the users.
The low prediction accuracy of both model formulations, coupled with the lack of data in the extremes of the applicability limits of the models, may suggest that the models' applicability limits should be revisited.
Based on the results, we suggest that the Standards applicability limits could be reduced to the following range \qty{18}{\celsius}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius} and \qty{10}{\celsius}~$\leq$~\ac{tr}~$\leq$~\qty{30}{\celsius} (instead of \qty{10}{\celsius}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius} and \qty{10}{\celsius}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius}),  until more data are collected.

Grouping the results into discrete categories introduces rounding errors since some participants reported \ac{tsv} on a continuous scale and the \ac{pmv} value is a continuous variable.
To compensate for this we plotted the \ac{pmv} and \ac{pmv-ce} values as a function of \ac{tsv} in Figure~\ref{fig:bubble_models_vs_tsv} and we plotted a \ac{lowess} curve.
% todo I would not use grey for the data, it is too similar to the gridlines. A light color is fine, light brown (depending on the color pallet used in the rest of the paper) could work. The regression line could be done even thicker. Be more explicit in the graph of what is measured and what is calculated. The Y axis should be labeled as PMV

\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bubble_models_vs_tsv}
    \caption{The \ac{lowess} curve shows the relationship between the raw \ac{pmv} and \ac{tsv} data.
    Raw data were then binned and a bubble chart is superimposed over the regression curve to aid the visualization of a large dataset.}
    \label{fig:bubble_models_vs_tsv}
\end{figure*}
The curve, in the Figure, is calculated using the individual data and not the binned data.
We binned the data to aid the visualization of a large dataset.
If the model is to accurately predict the thermal sensation of people, the regression line should pass through the origin of the cartesian plane and have a slope of 1.
The intercepts for the \ac{pmv} is \qty{-0.23} and for the \ac{pmv-ce} is \qty{-0.24} which means that the model has a bias of less than one-quarter of a thermal sensation interval. 
The intercept is close to 0 and this is aligned with the results shown above that the model can predict people who reported to be neutral \qty{55}{\percent} and \qty{58}{\percent} of the times.
Both \ac{pmv} formulations always underpredicted the thermal sensation of people who reported to be on the extremes of the \ac{tsv}; the slope of the curve was significantly lower than 1.
We discuss in detail in sub-Section~\ref{subsec:sources-of-error} which factors may be the underlying cause of these errors.

The intended aim of the \ac{pmv} model is not to accurately predict each individual thermal response from participants.
The \ac{pmv} model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.
Consequently, while the above-mentioned analysis is needed to compare for the same sample size in each thermal sensation group the \ac{pmv} and \ac{pmv-ce} accuracies, it focuses primarily on comparing individual \ac{tsv} to \ac{pmv}. 
This type of analysis does not conclusively disprove the inefficacy of the \ac{pmv} models in predicting the average thermal sensation of a large group of occupants.
This was already done in \mycite{Cheung2019}, where they grouped people based on the same thermal environment (same \ac{pmv} ranges), not the same thermal sensation vote as we did here, and then quantified how many times the \ac{pmv} model correctly predicted \ac{tsv}. 
They found that the accuracy of \ac{pmv} in predicting \ac{tsv} was only \qty{34}{\percent} \mycite{Cheung2019}.

\subsection{Model Bias}\label{sec:model-bias}
\subsubsection{Model Overall Bias}\label{subsec:model-overall-bias}
We calculated the overall bias of the model as previously done by \mycite{Humphreys2002} by subtracting the \ac{pmv} and  \ac{pmv-ce} models prediction from the self-reported \ac{tsv}.
Results are presented in Figure~\ref{fig:hist_discrepancies}.
% todo You need to check of the four distibution are normally distributed, if so, report mean and SD, if not report median and IQR, no need to report all of them. I guess that Pswetest is an indication of this, so, for the entire dataset it seems that it is not sufficiently normally distributed, so report only median and IQR. The Pswetest This index is not reported, explained, and interpreted in the paper. I would like to have these figure a bit longer in the Y-axis. they are kind of squized. The word "delta" does not mean difference. So, I suggest to use "diffference" or just show the equation Bias=PMVi-TSVi. 1) Can this chart be transformed in a letter value plot? The Caption is missing
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/hist_discrepancies}
    \caption{}
    \label{fig:hist_discrepancies}
\end{figure*} 
Overall both the \ac{pmv} and \ac{pmv-ce} models have a median bias lower than half of a thermal sensation interval.
Figure~\ref{fig:hist_discrepancies} contains the summary statistics, and the results highlight that the \ac{pmv-ce} had a higher bias than the \ac{pmv} when using the entire dataset.
The two models only differ when \ac{v} exceeds \qty{0.1}{\m\per\s}; we also calculated the bias of the model, including data points with \ac{v} higher than \qty{0.2}{\m\per\s} since above this value the differences in the outputs of the two models are more pronounced.
We set the limit to \qty{0.2}{\m\per\s} and not a higher value since, as previously shown in Figure~\ref{fig:dist_input_data}, we could only reliably test the accuracy of the models up to \ac{v} of \var{v_95_perc_max}.
We did not use \qty{0.1}{\m\per\s} as a threshold because the original difference between the two models was \qty{0.2}{\m\per\s}.
The effect of air movement is perceptible only when the airspeed is sufficiently high to disrupt the body's thermal plume (xxx).
The threshold in the standard was reduced to \qty{0.1}{\m\per\s} to simplify the usability and achieve one threshold for all the air speeds.
The results show that the median bias of the \ac{pmv} model reduced to \var{bias_median_pmv_0.2} from \var{bias_median_pmv_0}, meaning that at higher speed the \ac{pmv} model improved.
Surprisingly, we found that the median bias for the \ac{pmv-ce} increased to \var{bias_median_pmv_ce_0.2} from \var{bias_median_pmv_ce_0}.
These are worrying results since the \ac{pmv-ce} model claims to have a higher accuracy at higher values of \ac{v}, but this is not what was found from this data.
As far as we could find, there is no published paper proving a higher accuracy of the \ac{pmv-ce} compared to the \ac{pmv} model. 
The \ac{pmv-ce} has lower accuracy while making the code more complex and computationally intensive to run.

\subsubsection{Model Bias as Function of Four Input Variables}\label{subsec:model-bias-variable}
In the previous section, we presented the overall bias of the two \ac{pmv} formulations.
We then went on to determine how the bias varied as a function of \ac{tdb}, \ac{v}, \ac{met}, and \ac{clo}.
Figure~\ref{fig:dist_input_data} shows the results.
In each sub-figure, we have binned the results into discrete intervals of equal width, and the x-axis tick label is the center value of the interval. 
The width of the violin plot is not proportional to the number of points in that specific bin.
% todo This figure needs a lot of improvements. 1) I would use the letter-value (or boxplot), they may be easier to interpret, provide less info and could be aligned with the first figure (in particular if  The actual density distribution it is not adding much value. 2) The color for PMV and PMV_CE must be the same in the entire paper. 3) The y-label should be moved to the x-label and the y-label should explicit. I would remove the decimal values and show only the five integers -2, -1, 0, 1 and 2. 4) How is possible to have a velocity "centered" at 0? I would remove the red lines. If you want to underline the -0.5 and 0.5 range, then we should use the same strategy also for figure 7. The fact that you change graph time in the paper, it makes more difficult to interpret without adding a substantial amount of value. 
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/bias_models}
    \caption{}
    \label{fig:bias_models}
\end{figure*}
The results depict that the use of the \ac{pmv-ce} model did not improve the bias of the \ac{pmv} results across all four input variables.
On the contrary, the bias of the \ac{pmv-ce} diverged more from 0 as the value of \ac{v} increased.
These results are in agreement with those presented in the previous section and depict that the correction applied by the \ac{pmv-ce} model worsens the results.

\subsection{Comparison between the PMV formulations with the two-node model}\label{subsec:comparison-between-the-pmv-formulations-wih-the-two-node-model}
We compared how the results of the \ac{pmv} and \ac{pmv-ce} perform against the results of the two-node model because xxxx.
Figure~\ref{fig:pmv_two_node_comparison} was generated by calculating the three \ac{pmv} values using the same combination of \num{10000} randomly generated sets inputs.
The \ac{pmv} and \ac{pmv-ce} were then plotted on the y-axis while the \ac{pmvg} value was on the x-axis.
The Figure also shows the \ac{lowess} curves that are used to visualize the prevailing data trend.
The results of the \ac{pmv} model had a higher agreement than those obtained with the \ac{pmv-ce} model for \ac{pmvg} $\geq$ \qty{0.5}{\m\per\s}.
These results show the \ac{pmv-ce} is less similar to the \ac{pmvg} model than the \ac{pmv}, despite using it in the backend to calculate the cooling effect.
This is also an unexpected result and we will discuss why this may be in the following paragraph. 
% todo When a figure report how well a variable is predicted, for example x-measured vs x-predicted, the graph must be a square, otherwise it is visually skewed. 1) I suggest to use the words instead of numbers for PMV 2) We should be consistent with the PMV naming we used (do not use ASHRAE and ISO here) 3) why both models are far from Gagge for PMV<0, both model in general produce a cooler feeling than PMVgagge, we should explain why if we keep this paper.

\begin{figure*}[htb!]
    \centering
    \includegraphics[width=\textwidth]{figures/pmv_two_node_comparison}
    \caption{}
    \label{fig:pmv_two_node_comparison}
\end{figure*}

\subsection{Sources of Error}\label{subsec:sources-of-error}
In this Section, we identify some of the main reasons that could explain the difference in the results from the two \ac{pmv} models when compared against the \ac{tsv} available in the \ac{db2}.
Moreover, we tried to explain why we observed that the \ac{pmv-ce} model has a lower accuracy than the \ac{pmv} in predicting thermal sensation.
The issues are not reported in order of importance and more research is needed to determine how to improve the overall accuracy of the \ac{pmv} model.

\paragraph{Lack of Data}
T.

\paragraph{Heat Balance Equation and Calculation of the PMV value}
The \ac{pmv} model uses simplified heat balance equations to estimate the heat loss and gains from the human body to its surrounding environment.
This is a possible source of error since the model considers the human body to be a cylinder with constant width all uniformly covered with clothing where the heat is all generated in its core.
This is a great simplification since it ignores, for example, the fact that different parts of the body may not be covered by clothing, the metabolic heat generation is not uniform, the mass-to-surface ratios are not constant, and it ignores vasoconstriction and vasodilation.
However, the \ac{pmv} model erroneously assumes that the human body, aside from some edge cases, is either losing or gaining heat from its surrounding environment.
This is not the case, since the human body in most conditions observed indoors activates control strategies to maintain a constant core temperature.
Moreover, the \ac{pmv} model has been developed based on the assumption of steady-state heat transfer, however, this never precisely occurs since the human thermoregulatory system is always actively engaged to ensure a stable core temperature.

The \ac{pmv} model uses the overall heat losses or gains, which are only a proxy to the thermal stress that the human body has to undergo to maintain a constant internal temperature, to a \ac{pmv} value which should represent a \ac{tsv} vote.
Fanger created this equation based on a few experimental data collected in a laboratory, and despite the growing body of literature, this equation was never updated~\cite{Fanger1970}.
In the \ac{pmv} model the heat losses and gains are mapped using constant coefficients to a thermal sensation vote, disregarding the fact that the human body does not have the same ability to compensate for warm and cold conditions.

It is, therefore, not surprising to observe that the \ac{pmv} model has an `acceptable' accuracy in determining when people are `neutral`.
This is equivalent to no heat losses or gains (i.e., \ac{pmv}=0) which means that the body is dissipating all the internal metabolic heat production through conduction, convection, or radiation without needing to thermoregulate using sweating, shivering, vasoconstriction, vasodilation, behavioral adjustments, and non-shivering thermogenesis.
On the other hand, the model cannot correctly predict the thermal sensation of people who report being outside the three central thermal sensation votes (slightly cool, neutral and slightly warm).
We showed that people tend to be closer to neutrality than what the models predicted (see figure XXX), this is an indication that they were exposed to compensable thermal conditions. Therefore, the models' inability to account for these regulatory mechanisms is a fundamental limitation.

To continue using the \ac{pmv} model beyond an absolute value of 1, more research is needed to determine whether the inaccuracy of the model is related to either or a combination of the following factors and if those sources of inaccuracy can be reduced: 
i) the assumption that the heat losses/gains estimated by the \ac{pmv} are correlated to the thermoregulatory response that the body has to undertake to compensate for unsatisfactory thermal comfort conditions;
ii) the correlation coefficients used by Fanger to link the heat losses/gains and thermal stress are constant and not xxx;
iii) the regression coefficient between heat losses and heat gains and thermal stress are not the same for the heat losses and gains and two different equations should be used;
iv) the anchors used by the \ac{pmv} model are different from those used by participants in reporting their thermal sensation.
For example, while the \ac{pmv} model may assume that a score of 3 (hot) signifies the change between a compensable set of conditions to an un-compensable one, e.g., their skin wettedness reaches the max value ($w_{max}$), as suggested by Gagge's in his manuscript~\cite{GaggeSET}.
Participants in commercial buildings may, on the contrary, deem the environment to be `Hot' as soon as they start sweating and their \ac{w} crosses a critical threshold which is only a fraction of $w_{max}$.

It is likely that a combination of all the above-mentioned factors plays a role in explaining the discrepancies between the \ac{pmv} model and the self-reported \ac{tsv}.
The fourth hypothesis has gained significant momentum in the literature which led to the development of several \ac{pmv} formulations that apply correction factors to the final \ac{pmv} result without changing the underlying equation~\cite{Yao2022, Toftum2002}

\paragraph{Measurement Errors}
All the data contained in the \ac{db2} have been published in peer-reviewed manuscripts, however, measurement errors are inevitably present in the dataset.
This is due to the fact that all measurement instrumentation is not always calibrated before use, the uncertainty of the measurements may be large, the placement of the instrumentation may not be in the near proximity of the participant, and estimating clothing and metabolic rates from tables is a complex and prone to error task.
In an ideal scenario, errors would be random and would cancel out, not significantly affecting the overall bias of the model but only affecting the standard deviation of the error~\cite{Humphreys2002}.
However, this may not always be the case, for example, if a researcher often underestimated the clothing insulation of participants or their activity levels.

\paragraph{Thermal Sensation Scale}
One minor but not negligible source of error is using a discrete scale to assess thermal sensation (most of the \ac{tsv} in the \ac{db2} were collected using a discrete scale) while the \ac{pmv} output is continuous.
For example, even if the model was 100~\% accurate and the estimated \ac{pmv} value was \num{2.5} the person could only report being `warm (+2)' or `hot (+3)'.
Using a continuous, may in principle be a more robust approach and align with the output but \mycite{schweiker2019scales} showed that the distances between the anchors are not constant ~\cite{schweiker2019scales, schweiker2020evaluating}. 
They also showed that the \ac{tsv}'s assumption of its independence from contextual factors such as climate, season and language is flawed~\cite{schweiker2019scales, schweiker2020evaluating}.
An alternative solution would be to transform the \ac{pmv} output to an ordinal and categorical seven-point scale.

% \paragraph{Other neglected factors}
% Errors are also caused by the fact that the \ac{pmv} model is an approximation of a highly complex system (the human body and mind) that interacts with the surrounding environment.
% Fanger could not, therefore, include all the factors that contribute to the comfort vote and some of them have been omitted.

% todo write a paragraph trying to find the reason why the PMV CE model is less accurate.
% todo write a paragraph about what we recommend for ASHRAE 55

### 2023-08-29 Manuscript/src/sections/titleAbstract.tex
paper: reviewing Stefano's comments

    \title{PMV formulations: which one is the most accurate?}

@@

    \begin{abstract}
        The primary objectives of buildings is to shelter pople from the elements and provide a comfortable thermal environment.
        The \ac{pmv} model was therefore developed by Fanger more than 50 years ago to predict the thermal sensation of a large group of occupants sharing the same thermal environment.
        This should enable to determine the optimal temperature range to be maintained indoors to satisfy the great majority of the occupants.
        The \ac{pmv} has seen widespread adoption and is and has been used by practitioners to design and operate buildings and by scientists to predict thermal sensation.
        It is still currently used in its original form in the \gls{7730} despite the fact that several research studies have highlighted that the \ac{pmv} model has a low prediction accuracy, especially when it tries to predict the thermal sensation of people who are not thermally neutral.
        To compensate for that, several other formulations have been proposed, one of which one is currently used in the \gls{55}.
        Being the \gls{7730} and \gls{55} the most widely used Standards worldwide, this is a source of confusion for practitioners and researchers who find it difficult to decide which formulation to use.
        We aim to solve this problem by comparing the accuracy of the two main \ac{pmv} formulations, currently being used a reference in the above-mentioned standard, against \var{entries_db_used} thermal sensation votes collected in buildings.
        % We found that for V>0.1m/s the \gls{pmv-ce} is significantly more accurate (48%) than the PMV (42%) in predicting thermal neutrality.
        % In 95% of the cases, the error was less than one point.
        % The PMV and PMVCE accuracy were 35% and 36% for |PMV|≤2 and 43% and 44% for |PMV|≤1, respectively.
        % Neither PMV model could accurately predict ‘cold’, ‘cool’, ‘warm’, or ‘hot’ thermal sensations.
        % Consequently, the PMV should only be used in conditions close to neutrality within the range |PMV|≤1.
         We open-sourced the source code we used to calculate the \ac{pmv} and we used to analyse the data.
    \end{abstract}

    \begin{keyword}
        Personal thermal comfort model \sep Ecological momentary assessment \sep Skin temperature \sep Machine learning \sep Internet of Things (IoT)
    \end{keyword}

>>>>

    \title{PMV: Which formulation should you use the ISO 7730:2005 or ASHRAE 55:2020?}

@@

    \begin{abstract}
        The \ac{pmv} model was developed by Fanger more than 50 years ago to predict the mean thermal sensation of a group of people. 
        It is based on a heat balance model of the human body.

        \gls{55} states that the model can be used to determine ``the combinations of indoor thermal environmental factors and personal factors that will produce satisfactory thermal environmental conditions for a majority of the occupants within the space''.

        The \ac{pmv} has seen widespread adoption and has been used by professionals to design and operate buildings and by researchers to predict thermal sensation.
        It is still currently used in its original form in the \gls{7730} although it has a low prediction accuracy, especially outside thermal neutrality.
        Several other formulations have been proposed to compensate for its limitations, one of which is currently used in the \gls{55}.
        \gls{7730} and \gls{55} are the most widely used Standards worldwide, and their differences are a source of confusion for users who find it difficult to decide which model to use.

        We aim to solve this problem by comparing the accuracy of the two main \ac{pmv} formulations, against \var{entries_db_used} thermal sensation votes collected in buildings.
        We found that the \ac{pmv-ce} has the same accuracy as the original \ac{pmv} in predicting thermal neutrality, despite its increased complexity and computational resources needed to calculate the results.
        Surprisingly, the bias of the \ac{pmv-ce} model was greater (a higher number is negative) than the \ac{pmv} when we only analyzed the entries with \ac{v}$\geq$\qty{0.1}{\m\per\sec}.

        The \ac{pmv} and \ac{pmv-ce} accuracy were XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 2$ and XX\% and YY\% for $\mid$\ac{pmv}$\mid \leq 1$, respectively.
        Neither PMV model could accurately predict `cold’, `cool’, `warm’, or `hot’ thermal sensations.
        Consequently, the PMV should only be used in conditions close to neutrality within the range $\mid$\ac{pmv}$\mid \leq 1$, and we suggest reducing the applicability limit of the Standards to this range.
    \end{abstract}

    \begin{keyword}
        Thermal comfort model \sep ISO 7730 \sep ASHRAE 55 \sep Thermal sensation \sep Accuracy
    \end{keyword}
